Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2025-01-12 paper/main.tex
some style and phrasing clean up

Our primary research goal is to advance our understanding of what it means to
acquire deep, real-world conceptual knowledge. Traditional laboratory
approaches to studying learning and memory (e.g., list-learning studies) often
draw little distinction between memorization and understanding. Instead, these
studies typically focus on whether information is effectively encoded or
retrieved, rather than whether the information is \textit{understood}.
Approaches to studying conceptual learning, such as category learning
experiments, can begin to investigate the distinction between memorization and
understanding, often by training participants to distinguish arbitrary or
random features in otherwise meaningless categorized stimuli~\citep{ReilEtal82,
Este86a, Este86b, GlucEtal02, AshbMadd05, HulbNorm15}. However the objective of
real-world training, or learning from life experiences more generally, is often
to develop new knowledge that may be applied in \textit{useful} ways in the
future. In this sense, the gap between modern learning theories and modern
pedagogical approaches that inform classroom learning strategies is enormous:
most of our theories about \textit{how} people learn are inspired by
experimental paradigms and models that have only peripheral relevance to the
kinds of learning that students and teachers actually seek~\citep{Macl05,
HallGree08}. To help bridge this gap, our study uses course materials from real
online courses to inform, fit, and test models of real-world conceptual
learning. We show that these models recover meaningful relationships between
concepts presented during course lectures and tested by assessments, and that
these relationships can be leveraged to predict students' success on individual
quiz questions. We also provide a demonstration of how our models can be used to
construct ``maps'' of what students know, and how their knowledge changes with
training. In addition to helping to visually capture knowledge (and changes in
knowledge), we hope that such maps might lead to real-world tools for improving
how we educate. Taken together, our work shows that existing course materials
and evaluative tools like short multiple-choice quizzes may be leveraged to
gain highly detailed insights into what students know and how they learn.

@@

If we are able to accurately estimate a participant's knowledge about the
content tested by a given question, our estimates of their knowledge should
carry some predictive information about whether they are likely to answer that
question correctly or incorrectly. We developed a statistical approach to test
this claim. For each quiz question a participant answered, in turn, we used
Equation~\ref{eqn:prop} to estimate their knowledge at the given question's
embedding space coordinate based on other questions that participant answered
on the same quiz. We repeated this for all participants, and for each of the
three quizzes. Then, separately for each quiz, we fit a generalized linear
mixed model (GLMM) with a logistic link function to explain the probability of
correctly answering a question as a function of estimated knowledge for its
embedding coordinate, while accounting for random variation among participants
and questions (see \nameref{subsec:glmm}). To assess the predictive value of
the knowledge estimates, we compared each GLMM to an analogous (i.e., nested)
``null'' model that did not consider estimated knowledge using parametric
bootstrap likelihood-ratio tests.

\begin{figure}[tp]
    \centering
    \includegraphics[width=0.75\textwidth]{figs/predict-knowledge-questions}
    \caption{\textbf{Predicting success on held-out questions using estimated
    knowledge.} We used generalized linear mixed models (GLMMs) to model the
    likelihood of correctly answering a quiz question as a function of
    estimated knowledge for its embedding coordinate (see
    \nameref{subsec:glmm}). Separately for each quiz (column), we examined this
    relationship based on three different sets of knowledge estimates:
    knowledge for each question based on all other questions the same
    participant answered on the same quiz (``All questions''; top row),
    knowledge for each question about one lecture based on all other questions
    (from the same participant and quiz) about the \textit{same} lecture
    (``Within-lecture''; middle rows), and knowledge for each question about
    one lecture based on all questions (from the same participant and quiz)
    about the \textit{other} lecture (``Across-lecture''; bottom rows). The
    backgrounds in each panel display kernel density estimates of the relative
    observed proportions of correctly (blue) versus incorrectly (red) answered
    questions, for each level of estimated knowledge along the $x$-axis. The
    black curves display the (population-level) GLMM-predicted probabilities of
    correctly answering a question as a function of estimated knowledge. Error
    ribbons denote 95\% confidence intervals.}

@@

We carried out three different versions of the analyses described above,
wherein we considered different sources of information in our estimates of
participants' knowledge for each quiz question. First, we estimated knowledge
at each question's embedding coordinate using \textit{all} other questions
answered by the same participant on the same quiz (``All questions'';
Fig.~\ref{fig:predictions}, top row). This test was intended to assess the
overall predictive power of our approach. Second, we estimated knowledge for
each question about a given lecture using only the other questions (from the
same participant and quiz) about that \textit{same} lecture
(``Within-lecture''; Fig.~\ref{fig:predictions}, middle rows). This test was
intended to assess the \textit{specificity} of our approach by asking whether
our predictions could distinguish between questions about different content
covered by the same lecture. Third, we estimated knowledge for each question
about one lecture using only the questions (from the same participant and quiz)
about the \textit{other} lecture (``Across-lecture'';
Fig.~\ref{fig:predictions}, bottom rows). This test was intended to assess the
\textit{generalizability} of our approach by asking whether our predictions
could extend across the content areas of the two lectures. When computing these
knowledge estimates, we used a rebalancing procedure to ensure that (for a given
participant and quiz) the knowledge estimates for correctly and incorrectly
answered questions were computed from the same proportion of correctly answered
questions (see~\nameref{subsec:glmm}).

When we fit a GLMM to estimates of participants' knowledge for each Quiz~1
question based on all other Quiz~1 questions, we found that higher estimated
knowledge for a given question predicted a greater likelihood of answering it
correctly (odds ratio $(OR) = 8.126$, 95\%\ $\textnormal{CI} = [3.116,\
20.123]$, likelihood-ratio test statistic $(\lambda_{LR}) = 17.002,\ p <
0.001$). This relationship held when we repeated this analysis for Quiz~2 ($OR
= 14.902,\ 95\%\ \textnormal{CI} = [4.976,\ 39.807],\ \lambda_{LR} = 25.408,\
p < 0.001$) and again for Quiz~3 ($OR = 37.409,\ 95\%\ \textnormal{CI} =
[10.425,\ 107.145],\ \lambda_{LR} = 40.948,\ p < 0.001$). Taken together, these
results suggest that our knowledge estimates can reliably predict participants'
performance on individual questions when aggregated across all quiz content.

We observed a similar set of results when we restricted our estimates of
participants' knowledge for questions about each lecture to consider only
their performance on other questions about the \textit{same} lecture.
Specifically, for Quiz~1, participants' knowledge of \textit{Four Fundamental
Forces}-related questions, estimated from their performance on other
\textit{Four Fundamental Forces}-related questions, was predictive of their
ability to answer those questions correctly ($OR = 15.934,\ 95\%\
\textnormal{CI} = [5.173,\ 38.005],\ \lambda_{LR} = 40.971,\ p = 0.001$). The
same was true of participants' estimated knowledge for \textit{Birth of
Stars}-related questions based on their performance on other \textit{Birth of
Stars}-related questions ($OR = 9.775,\ 95\%\ \textnormal{CI} = [2.93,\
25.08],\ \lambda_{LR} = 13.924,\ p = 0.001$). These within-lecture knowledge
estimates also predicted success on questions about both lectures when we
computed them analogously for Quiz~2 (\textit{Four Fundamental Forces}: $OR =
35.126,\ 95\%\ \textnormal{CI} = [5.113,\ 123.868],\ \lambda_{LR} = 32.251,\
p < 0.001$; \textit{Birth of Stars}: $OR = 4.717,\ 95\%\ \textnormal{CI} =
[2.021,\ 9.844],\ \lambda_{LR} = 16.788,\ p < 0.001$). For Quiz~3, we found that
within-lecture knowledge estimates predicted participants' success on
\textit{Birth of Stars}-related questions ($OR = 16.902,\ 95\%\ \textnormal{CI}
= [3.353,\ 53.265],\ \lambda_{LR} = 23.233,\ p < 0.001$) but not on \textit{Four
Fundamental Forces}-related questions ($OR = 2.485,\ 95\%\ \textnormal{CI} =
[0.724,\ 8.366],\ \lambda_{LR} = 1.984,\ p = 0.170$). This may indicate that the
within-lecture knowledge estimates are susceptible to ceiling effects in
participants' quiz performance. On Quiz~3, after viewing both lectures, no
participant answered more than three \textit{Four Fundamental Forces}-related
questions incorrectly, and all but five participants (out of 50) answered two or
fewer questions incorrectly. (This was the only subset of questions about either
lecture, across all three quizzes, for which this was true.) Because of this,
when we held out one incorrectly answered \textit{Four Fundamental Forces}
question from a given participant's Quiz~3 responses and estimated their
knowledge at its embedding coordinate using the remaining \textit{Four
Fundamental Forces} questions they answered, for 90\% of participants, that
estimate leveraged information about at most a single other question they were
\textit{not} able to correctly answer. This broad homogeneity in participants'
success on questions used to estimate their knowledge may have hurt our ability
to accurately characterize the specific (and by Quiz~3, relatively few) aspects
of the lecture content they did \textit{not} know about. Taken together, these
results suggest that our knowledge estimates can reliably distinguish between
questions about different content covered by a single lecture, provided there is
sufficient diversity in participants' quiz responses to extract meaningful
information about both what they know and what they do not know.

@@

That the knowledge predictions derived from the text embedding space reliably
distinguish between held-out correctly versus incorrectly answered questions
(Fig.~\ref{fig:predictions}) suggests that spatial relationships within this
space can help explain what participants know. But how far does this
explanatory power extend? For example, suppose we know that a participant
correctly answered a question at embedding coordinate $x$. As we move farther
away from $x$ in the embedding space, how does the likelihood that the
participant knows about the content at a given location ``fall off'' with
distance? Conversely, suppose the participant instead answered that same
question \textit{in}correctly. Again, as we move farther away from $x$ in the
embedding space, how does the likelihood that the participant does \textit{not}
know about a coordinate's content change with distance? We reasoned that,
assuming our embedding space is capturing something about how individuals
actually organize their knowledge, a participant's ability to answer questions
embedded very close to $x$ should tend to be similar to their ability to answer
the question embedded \textit{at} $x$. Whereas at another extreme, once we
reach some sufficiently large distance from $x$, our ability to infer whether
or not a participant will correctly answer a question based on their ability to
answer the question at $x$ should be no better than guessing based on their
\textit{overall} proportion of correctly answered questions. In other words,
beyond the maximum distance at which the participant's ability to answer the
question at $x$ is informative of their ability to answer a second question at
location $y$, then guessing the outcome at $y$ based on $x$ should be no more
successful than guessing based on a measure that does not consider embedding
space distance.

@@

All of the statistical tests performed in our study were two-sided. The 95\%
confidence intervals we reported for each correlation were estimated by
generating 10,000 bootstrap distributions of correlation coefficients by
sampling (with replacement) from the observed data.

@@

We adapted an approach we developed in prior work~\citep{HeusEtal21} to embed
each moment of the two lectures and each question in our pool in a common
representational space. Briefly, our approach uses a topic model~\citep[Latent
Dirichlet Allocation; ][]{BleiEtal03} trained on a set of documents, to
discover a set of $k$ ``topics'' or ``themes.'' Formally, each topic is
defined as a distribution of weights over words in the model's vocabulary
(i.e., the union of all unique words, across all documents, excluding ``stop
words.''). Conceptually, each topic is intended to give larger weights to words
that are semantically related (as inferred from their tendency to co-occur in
the same document). After fitting a topic model, each document in the training
set, or any \textit{new} document that contains at least some of the words in
the model's vocabulary, may be represented as a $k$-dimensional vector
describing how much the document (most probably) reflects each topic. To select
an appropriate $k$ for our model, as a starting point, we identified the
minimum number of topics that yielded at least one ``unused'' topic (i.e., in
which all words in the vocabulary were assigned uniform weights) after
training. This indicated that the number of topics was sufficient to capture
the set of latent themes present in the two lectures (from which we constructed
our document corpus, as described below). We found this value to be $k = 15$
topics. We found that with a limited number of additional adjustments
following~\citet{BoydEtal14}, such as removing corpus-specific stop-words, the
model yielded (subjectively) sensible and coherent topics. The distribution of
weights over words in the vocabulary for each discovered topic is shown in
Supplementary Figure~\topicWordWeights, and each topic's top-weighted words may
be found in Supplementary Table~\topics.

@@

\begin{equation}
    \mathrm{ncorr}(x, y) = \frac{\mathrm{corr}(x, y) - \mathrm{mincorr}}{\mathrm{maxcorr} - \mathrm{mincorr}},
\end{equation}
and where $\mathrm{mincorr}$ and $\mathrm{maxcorr}$ are the minimum and maximum
correlations between any lecture timepoint and question, taken over all
timepoints in the given lecture, and all five questions \textit{about} that
lecture appearing on the given quiz. We also define $f(s, \Omega)$ as the
$s$\textsuperscript{th} topic vector from the set of topic vectors $\Omega$.
Here $t$ indexes the set of lecture topic vectors $L$, and $i$ and $j$ index
the topic vectors of questions $Q$ used to estimate the knowledge trace. Note
that ``correct'' denotes the set of indices of the questions the participant
answered correctly on the given quiz.

Intuitively, $\mathrm{ncorr}(x, y)$ is the correlation between two topic
vectors (e.g., the topic vector $x$ for one timepoint in a lecture and the
topic vector $y$ for one question), normalized by the minimum and maximum
correlations (across all timepoints $t$ and questions $j$) to range between 0
and 1, inclusive. Equation~\ref{eqn:prop} then computes the weighted average
proportion of correctly answered questions about the content presented at
timepoint $t$, where the weights are given by the normalized correlations
between timepoint $t$'s topic vector and the topic vectors for each question.
The normalization step (i.e., using $\mathrm{ncorr}$ instead of the raw
correlations) ensures that every question contributes some non-negative amount
to the knowledge estimate.

>>>>

Our primary research goal is to advance our understanding of what it means to
acquire deep, real-world conceptual knowledge. Traditional laboratory
approaches to studying learning and memory (e.g., list-learning studies) often
draw little distinction between memorization and understanding. Instead, these
studies typically focus on whether information is effectively encoded or
retrieved, rather than whether the information is \textit{understood}.
Approaches to studying conceptual learning, such as category learning
experiments, can begin to investigate the distinction between memorization and
understanding, often by training participants to distinguish arbitrary or
random features in otherwise meaningless categorized stimuli~\citep{ReilEtal82,
Este86a, Este86b, GlucEtal02, AshbMadd05, HulbNorm15}. However, the objective of
real-world training, or learning from life experiences more generally, is often
to develop new knowledge that may be applied in \textit{useful} ways in the
future. In this sense, the gap between modern learning theories and modern
pedagogical approaches that inform classroom learning strategies is enormous:
most of our theories about \textit{how} people learn are inspired by
experimental paradigms and models that have only peripheral relevance to the
kinds of learning that students and teachers actually seek~\citep{Macl05,
HallGree08}. To help bridge this gap, our study uses course materials from real
online courses to inform, fit, and test models of real-world conceptual
learning. We show that these models recover meaningful relationships between
concepts presented during course lectures and tested by assessments, and that
these relationships can be leveraged to predict students' success on individual
quiz questions. We also provide a demonstration of how our models can be used to
construct ``maps'' of what students know, and how their knowledge changes with
training. In addition to helping to visually capture knowledge (and changes in
knowledge), we hope that such maps might lead to real-world tools for improving
how we educate. Taken together, our work shows that existing course materials
and evaluative tools like short multiple-choice quizzes may be leveraged to
gain highly detailed insights into what students know and how they learn.

@@

If we are able to accurately estimate a participant's knowledge about the
content tested by a given question, our estimates of their knowledge should
carry some predictive information about whether they are likely to answer that
question correctly or incorrectly. We developed a statistical approach to test
this claim. For each quiz question a participant answered, in turn, we used
Equation~\ref{eqn:prop} to estimate their knowledge at the given question's
embedding-space coordinate based on other questions that participant answered
on the same quiz. We repeated this for all participants, and for each of the
three quizzes. Then, separately for each quiz, we fit a generalized linear
mixed model (GLMM) with a logistic link function to explain the probability of
correctly answering a question as a function of estimated knowledge for its
embedding coordinate, while accounting for varied effects of individual 
participants and questions (see \nameref{subsec:glmm}). To assess the predictive 
value of the knowledge estimates, we compared each GLMM to an analogous (i.e., 
nested) ``null'' model that assumed these estimates carried no predictive information using parametric bootstrap likelihood-ratio tests.

\begin{figure}[tp]
    \centering
    \includegraphics[width=0.75\textwidth]{figs/predict-knowledge-questions}
    \caption{\textbf{Predicting success on held-out questions using estimated
    knowledge.} We used generalized linear mixed models (GLMMs) to model the
    probability of correctly answering a quiz question as a function of
    estimated knowledge for its embedding coordinate (see
    \nameref{subsec:glmm}). Separately for each quiz (column), we examined this
    relationship based on three different sets of knowledge estimates:
    knowledge for each question based on all other questions the same
    participant answered on the same quiz (``All questions''; top row),
    knowledge for each question about one lecture based on all other questions
    (from the same participant and quiz) about the \textit{same} lecture
    (``Within-lecture''; middle rows), and knowledge for each question about
    one lecture based on all questions (from the same participant and quiz)
    about the \textit{other} lecture (``Across-lecture''; bottom rows). The
    backgrounds in each panel display kernel density estimates of the relative
    observed proportions of correctly (blue) versus incorrectly (red) answered
    questions, for each level of estimated knowledge along the $x$-axis. The
    black curves display the (population-level) GLMM-predicted probabilities of
    correctly answering a question as a function of estimated knowledge. Error
    ribbons denote 95\% confidence intervals.}

@@

We carried out three different versions of the analyses described above,
wherein we considered different sources of information in our estimates of
participants' knowledge for each quiz question. First, we estimated knowledge
at each question's embedding coordinate using \textit{all} other questions
answered by the same participant on the same quiz (``All questions'';
Fig.~\ref{fig:predictions}, top row). This test was intended to assess the
overall predictive power of our approach. Second, we estimated knowledge for
each question about a given lecture using only the other questions (from the
same participant and quiz) about that \textit{same} lecture
(``Within-lecture''; Fig.~\ref{fig:predictions}, middle rows). This test was
intended to assess the \textit{specificity} of our approach by asking whether
our predictions could distinguish between questions about different content
covered by the same lecture. Third, we estimated knowledge for each question
about one lecture using only the questions (from the same participant and quiz)
about the \textit{other} lecture (``Across-lecture'';
Fig.~\ref{fig:predictions}, bottom rows). This test was intended to assess the
\textit{generalizability} of our approach by asking whether our predictions
could extend across the content areas of the two lectures. When estimating
participants' knowledge, we used a rebalancing procedure to ensure that (for a 
given participant and quiz) their knowledge estimates for correctly and 
incorrectly answered questions were computed from the same underlying proportion 
of correctly answered questions (see~\nameref{subsec:glmm}).

When we fit a GLMM to estimates of participants' knowledge for each Quiz~1
question based on all other Quiz~1 questions, we found that higher estimated
knowledge for a given question predicted a greater likelihood of answering it
correctly (odds ratio $(OR) = 8.126$, 95\%\ $\textnormal{CI} = [3.116,\
20.123]$, likelihood-ratio test statistic $(\lambda_{LR}) = 17.002,\ p <
0.001$). This relationship held when we repeated this analysis for Quiz~2 ($OR
= 14.902,\ 95\%\ \textnormal{CI} = [4.976,\ 39.807],\ \lambda_{LR} = 25.408,\
p < 0.001$) and again for Quiz~3 ($OR = 37.409,\ 95\%\ \textnormal{CI} =
[10.425,\ 107.145],\ \lambda_{LR} = 40.948,\ p < 0.001$). Taken together, these
results suggest that our knowledge estimates can reliably predict participants'
performance on individual questions when they incorporate information from all 
(other) quiz content.

We observed a similar set of results when we restricted our estimates of
participants' knowledge for questions about each lecture to consider only
their performance on other questions about the \textit{same} lecture.
Specifically, for Quiz~1, participants' knowledge of \textit{Four Fundamental
Forces}-related questions, estimated from their performance on other
\textit{Four Fundamental Forces}-related questions, was predictive of their
ability to answer those questions correctly ($OR = 15.934,\ 95\%\
\textnormal{CI} = [5.173,\ 38.005],\ \lambda_{LR} = 40.971,\ p = 0.001$). The
same was true of participants' estimated knowledge for \textit{Birth of
Stars}-related questions based on their performance on other \textit{Birth of
Stars}-related questions ($OR = 9.775,\ 95\%\ \textnormal{CI} = [2.93,\
25.08],\ \lambda_{LR} = 13.924,\ p = 0.001$). These within-lecture knowledge
estimates also predicted success on questions about both lectures when we
computed them analogously for Quiz~2 (\textit{Four Fundamental Forces}: $OR =
35.126,\ 95\%\ \textnormal{CI} = [5.113,\ 123.868],\ \lambda_{LR} = 32.251,\
p < 0.001$; \textit{Birth of Stars}: $OR = 4.717,\ 95\%\ \textnormal{CI} =
[2.021,\ 9.844],\ \lambda_{LR} = 16.788,\ p < 0.001$). For Quiz~3, we found that
within-lecture knowledge estimates predicted participants' success on
\textit{Birth of Stars}-related questions ($OR = 16.902,\ 95\%\ \textnormal{CI}
= [3.353,\ 53.265],\ \lambda_{LR} = 23.233,\ p < 0.001$) but not on \textit{Four
Fundamental Forces}-related questions ($OR = 2.485,\ 95\%\ \textnormal{CI} =
[0.724,\ 8.366],\ \lambda_{LR} = 1.984,\ p = 0.170$). This may indicate that the
within-lecture knowledge estimates are susceptible to ceiling effects in
participants' quiz performance. On Quiz~3, after viewing both lectures, no
participant answered more than three \textit{Four Fundamental Forces}-related
questions incorrectly, and all but five participants (out of 50) answered two or
fewer questions incorrectly. (This was the only subset of questions about either
lecture, across all three quizzes, for which this was true.) Because of this,
when we held out one incorrectly answered 
\textit{Four Fundamental Forces}-related question from a given participant's 
Quiz~3 responses and estimated their knowledge at its embedding coordinate using 
the remaining \textit{Four Fundamental Forces}-related questions they answered, 
for 90\% of participants, that estimate leveraged information about at most a 
single other question they were \textit{not} able to correctly answer. This
broad homogeneity in participants' success on questions used to estimate their 
knowledge may have hurt our ability to accurately characterize the specific (and 
by Quiz~3, relatively few) aspects of the lecture content they did \textit{not} 
know about. Taken together, these results suggest that our knowledge estimates 
can reliably distinguish between questions about different content covered by a 
single lecture, provided there is sufficient diversity in participants' quiz 
responses to extract meaningful information about both what they know and what 
they do not know.

@@

That the knowledge predictions derived from the text embedding space reliably
distinguish between held-out correctly versus incorrectly answered questions
(Fig.~\ref{fig:predictions}) suggests that spatial relationships within this
space can help explain what participants know. But how far does this
explanatory power extend? For example, suppose we know that a participant
correctly answered a question at embedding coordinate $x$. As we move farther
away from $x$ in the embedding space, how does the likelihood that the
participant knows about the content at a given location ``fall off'' with
distance? Conversely, suppose the participant instead answered that same
question \textit{incorrectly}. Again, as we move farther away from $x$ in the
embedding space, how does the likelihood that the participant does \textit{not}
know about a coordinate's content change with distance? We reasoned that,
assuming our embedding space is capturing something about how individuals
actually organize their knowledge, a participant's ability to answer questions
embedded very close to $x$ should tend to be similar to their ability to answer
the question embedded \textit{at} $x$. Whereas at another extreme, once we
reach some sufficiently large distance from $x$, our ability to infer whether
or not a participant will correctly answer a question based on their ability to
answer the question at $x$ should be no better than guessing based on their
\textit{overall} proportion of correctly answered questions. In other words,
beyond the maximum distance at which the participant's ability to answer the
question at $x$ is informative of their ability to answer a second question at
location $y$, then guessing the outcome at $y$ based on $x$ should be no more
successful than guessing based on a measure that does not consider embedding
space distance.

@@

All of the statistical tests performed in our study were two-sided. The 95\%
confidence intervals we reported for each correlation were estimated from 
bootstrap distributions of 10,000 correlation coefficients obtained by
sampling (with replacement) from the observed data.

@@

We adapted an approach we developed in prior work~\citep{HeusEtal21} to embed
each moment of the two lectures and each question in our pool in a common
representational space. Briefly, our approach uses a topic model~\citep[Latent
Dirichlet Allocation; ][]{BleiEtal03} trained on a set of documents, to
discover a set of $k$ ``topics'' or ``themes.'' Formally, each topic is
defined as a distribution of weights over words in the model's vocabulary
(i.e., the union of all unique words, across all documents, excluding ``stop
words''). Conceptually, each topic is intended to give larger weights to words
that are semantically related (as inferred from their tendency to co-occur in
the same document). After fitting a topic model, each document in the training
set, or any \textit{new} document that contains at least some of the words in
the model's vocabulary, may be represented as a $k$-dimensional vector
describing how much the document (most probably) reflects each topic. To select
an appropriate $k$ for our model, as a starting point, we identified the
minimum number of topics that yielded at least one ``unused'' topic (i.e., in
which all words in the vocabulary were assigned uniform weights) after
training. This indicated that the number of topics was sufficient to capture
the set of latent themes present in the two lectures (from which we constructed
our document corpus, as described below). We found this value to be $k = 15$
topics. We found that with a limited number of additional adjustments
following~\citet{BoydEtal14}, such as removing corpus-specific stop-words, the
model yielded (subjectively) sensible and coherent topics. The distribution of
weights over words in the vocabulary for each discovered topic is shown in
Supplementary Figure~\topicWordWeights, and each topic's top-weighted words may
be found in Supplementary Table~\topics.

@@

\begin{equation}
    \mathrm{ncorr}(x, y) = \frac{\mathrm{corr}(x, y) - \mathrm{mincorr}}{\mathrm{maxcorr} - \mathrm{mincorr}},
\end{equation}
and where $\mathrm{mincorr}$ and $\mathrm{maxcorr}$ are the minimum and maximum
correlations between any lecture timepoint and question, taken over all
timepoints in the given lecture and all questions \textit{about} that
lecture appearing on the given quiz. We also define $f(s, \Omega)$ as the
$s$\textsuperscript{th} topic vector from the set of topic vectors $\Omega$.
Here $t$ indexes the set of lecture topic vectors $L$, and $i$ and $j$ index
the topic vectors of questions $Q$ used to estimate the knowledge trace. Note
that ``$\mathrm{correct}$'' denotes the set of indices of the questions the 
participant answered correctly on the given quiz.

Intuitively, $\mathrm{ncorr}(x, y)$ is the correlation between two topic
vectors (e.g., the topic vector $x$ for one timepoint in a lecture and the
topic vector $y$ for one question on a quiz), normalized by the minimum and 
maximum correlations (across all timepoints $t$ and questions $j$) to range 
between 0 and 1, inclusive. Equation~\ref{eqn:prop} then computes the weighted 
average proportion of correctly answered questions about the content presented 
at timepoint $t$, where the weights are given by the normalized correlations
between timepoint $t$'s topic vector and the topic vectors for each question.
The normalization step (i.e., using $\mathrm{ncorr}$ instead of the raw
correlations) ensures that every question contributes some non-negative amount
to the knowledge estimate.
